---
title: "Statistical Inference Course Project 2"
author: "Jackson Chou"
date: "May 29, 2016"
output: pdf_document
---

##Overview

This project analyzes the ToothGrowth data in the R datasets package.  I will be providing a basic summary of the dataset,
comparing tooth growth by supp and dose.

##Data

```{r}
library(datasets)
library(ggplot2)
library(gridExtra)
library(GGally)

#Loading the data
data(ToothGrowth)

```

##Summary of the ToothGrowth data
Let's take a look at the structure of our ToothGrowth dataset.

```{r}
str(ToothGrowth)
summary(ToothGrowth)
```

There are 60 observations with 3 variables: length (numeric), supplement (factor) and dosage levels (numeric).

```{r}
table(ToothGrowth$supp, ToothGrowth$dose)
```

In this dataset, there are two types of supplements, "VC" and "OJ".  Each supplement has three dosage levels: 0.5, 1.0, and 2.0.  Based on the table, there are ten observations for each combination of supplement and dosage level.

```{r, echo = FALSE}

p1 <- ggplot(data=ToothGrowth, aes(x=supp,y=len,fill=supp)) +
  geom_boxplot() + 
  theme(legend.position="none") + 
  facet_grid(.~dose) 

p2 <- ggplot(data=ToothGrowth, aes(x=supp,y=len,fill=supp)) +
  geom_boxplot()



grid.arrange( p1, p2, ncol = 1, nrow=2)


```

## Compare tooth growth by supp and dose
We'll use a two sample t-test to compare the difference of tooth growth by supplement and dosage levels.

Dosage level at 0.5:
```{r}
#Run t.test
t.test(len~supp, ToothGrowth[ToothGrowth$dose == 0.5, ])

```


Dosage level at 1:

```{r}
#Run t.test
t.test(len~supp, ToothGrowth[ToothGrowth$dose == 1, ])

```

Dosage level at 2:
```{r}
#Run t.test
t.test(len~supp, ToothGrowth[ToothGrowth$dose == 2, ])

```

##Conclusions
Based on the t-test analysis, it is statistically significant that OJ supplment results in longer tooth growth for dosage levels of 0.5 and 1.0 compared to VC supplements.  However, based on the p-value for dosage level 2, we cannot conclude that the OJ supplement results in better tooth growth compared to VC supplement.


